### 2025-05-08 writing/appendix-a-cleaning-telemetry-data.Rmd
wrote appendix for outlier cleaning (pdf > 100 MB)

---
title: \Large Predicting how climate change will affect how and where terrestrial mammals will move in British Columbia, Canada
subtitle: "Appendix A: Cleaning telemetry data"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [braes, biol]
  - name: Chris H. Fleming
    institute: [ucf, smith]
  - name: Siobhan Darlington
    institute: [braes, biol]
  - name: Adam T. Ford
    institute: [braes, biol]
  - name: Karen E. Hodges
    institute: [braes, biol]
  - name: Kirk Safford
    institute: bcp
  - name: Robert Serrouya
    institute: [braes, biol, biodiv]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - bcp: BC Parks
  - biodiv: Wildlife Science Centre, Biodiversity Pathways, University of British Columbia Okanagan, Revelstoke, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'bc-mammals-temperature.bib'
csl: 'freshwater-biology.csl'
# csl: 'nature.csl'
fontsize: 12pt
# indent: true
header-includes:
    - \renewcommand{\figurename}{Figure S\!\!} # for "Figure Ax.
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage{float} # to use fig.pos="H"
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    latex_engine: xelatex # for Delta symbol in table
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: true
---

```{r setup, include=FALSE}
# set chunk defaults
knitr::opts_chunk$set(echo = TRUE, out.width = '100%',
                      fig.align = 'center', cache = TRUE, warning = FALSE,
                      message = FALSE,
                      root.dir = '~/GitHub/bc-mammals-temperature')
```

\clearpage

\tableofcontents

\clearpage

# Setup

We start by attaching the necessary packages and sourcing some custom functions. The custom functions are all available on GitHub in the `functions` folder of the repository located at [https://github.com/QuantitativeEcologyLab/bc-mammals-temperature](https://github.com/QuantitativeEcologyLab/bc-mammals-temperature). The original cleaning script is located at [https://github.com/QuantitativeEcologyLab/bc-mammals-temperature/blob/main/analysis/02-remove-outliers.R](https://github.com/QuantitativeEcologyLab/bc-mammals-temperature/blob/main/analysis/02-remove-outliers.R). We suggest looking at the script since this appendix does not include the interactive maps that are generated with `plot_adj(..., map = TRUE)`, since they could not be included in the pdf output easily. However, the interacive maps often show linear features (e.g., roads) and other geographic features (e.g., mountain ridges, rivers) that were useful to see during the data cleaning. Therefore, we have left the function calls in the document for completeness, but we have commented them out to prevent them from being executed. Below is an example of a map with some locations for wolf BW008, as generated by running\newline`plot_adj('BW008', max_speed = 3.5, map = TRUE)`.

\singlespacing

```{r map-example, echo=FALSE, out.height='40%'}
setwd('writing')
knitr::include_graphics(path = '../figures/BW008-map-example.jpg')
```

\clearpage

```{r packages}
library('dplyr')     # for data wrangling (mutate(), %>%, etc.)
library('tidyr')     # for data wrangling (nest(), unnest(), pivot_*, etc.)
library('purrr')     # for functional programming (map_***(), etc.)
library('ctmm')      # for movement models
library('lubridate') # for working with dates
library('mapview')   # for interactive maps
library('ggplot2')   # for fancy plots
theme_set(theme_bw())

# source custom functions
source('functions/outlier_plots.R') # to plot outlier diagnostic plots
source('functions/check_animal.R') # to run diagnostic plots
source('functions/plot_adj.R') # to plot n adjacent locations
source('functions/flag_outlier.R') # to mark outliers
source('functions/remove_outlier_flags.R') # to start over with an animal
```

Before cleaning the telemetry data, we do some basic checks:

```{r checks}
# dataset checks ----
# some NA timestamps (not enough to make a difference)
readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(is.na(timestamp)) %>%
  group_by(dataset_name) %>%
  summarize(n = n())

# ensure all only have one column of HDOP
readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(! is.na(timestamp)) %>%
  mutate(outlier = if_else(outlier, 1, 0)) %>%
  filter(! is.na(hdop) + ! is.na(HDOP) > 1) %>%
  nrow()
```

<!-- to avoid splitting code chunk -->
\clearpage

```{r checks-2}
# should only have one column of DOP
readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(! is.na(timestamp)) %>%
  mutate(outlier = if_else(outlier, 1, 0)) %>%
  filter(! is.na(DOP) + ! is.na(gps.dop) > 1) %>%
  nrow()
```

We can now import the data...

```{r import-data}
# import the full dataset ----
# import the dataset after dropping NAs
d <- readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(! is.na(timestamp)) %>%
  mutate(hdop = if_else(is.na(hdop), HDOP, hdop), # merge error columns
         dop = if_else(is.na(DOP), gps.dop, DOP),
         original_outliers = outlier,
         outlier = if_else(outlier, 1, 0)) %>% # make numeric
  select(! c(HDOP, DOP, gps.dop)) %>% # remove duplicate columns
  relocate(dop, .after = pdop) %>%
  nest(tel = ! c(species, dataset_name, animal))
d # nested tibble of tracking datasets
```

... and do some last few check before cleaning.

```{r final-checks}
# ensure animal IDs are unique
sum(duplicated(d$animal))

# check which species have DOP values
d %>%
  unnest(tel) %>%
  group_by(species) %>%
  summarise(across(c(dop, hdop, pdop), ~ sum(.x, na.rm = TRUE))) %>%
  mutate(has_dop = dop + hdop + pdop > 0) %>%
  arrange(desc(has_dop))

#' see`finite`: some SM caribou have different values for hdop and dop
#' other datasets are ok
d %>%
  unnest(tel) %>%
  mutate(has_hdop = ! is.na(hdop),
         has_dop = ! is.na(dop),
         has_pdop = ! is.na(pdop)) %>%
  group_by(dataset_name, has_hdop, has_dop, has_pdop) %>%
  summarise(hdop = mean(hdop, na.rm = TRUE),
            pdop = mean(pdop, na.rm = TRUE),
            dop = mean(dop, na.rm = TRUE)) %>%
  mutate(finite = map_int(1:n(),
                          ~ sum(! is.nan(c(hdop[.x], pdop[.x], dop[.x])))))

d %>%
  filter(dataset_name == 'Rangifer_tarandus_southern_mountain') %>%
  unnest(tel) %>%
  filter(! is.na(hdop) & ! is.na(dop)) %>%
  plot(hdop ~ dop, ., main = paste('n =', nrow(.)))
abline(a = 0, b = 1, col = 'red', lwd = 2)
```

\clearpage

# *Canis lupus*

```{r canis-lupus, fig.pos='!H', fig.dim=c(10, 12)}
# BW008
out <- check_animal('BW008')
plot_adj('BW008', max_speed = 3.5)
# plot_adj('BW008', max_speed = 3.5, map = TRUE)
plot_adj('BW008', max_angle = 178, max_speed = 1) # using a linear feature
plot_adj('BW008', max_angle = 90, max_speed = 2) # realistic movement
plot_adj('BW008', max_angle = 170, max_speed = 0.5) # fine sampling
out[out$angle > 170 & out$speed > 0.5, ] # 15-minute sampling

# BW009
out <- check_animal('BW009')
plot_adj('BW009', max_angle = 179, max_speed = 1) # potential outlier
plot_adj('BW009', max_angle = 179.7, max_speed = 1, n_adj=200) # uses path
plot_adj('BW009', max_angle = 175, max_speed = 1, n_adj = 12) # ok
plot_adj('BW009', max_angle = 170, min_angle = 175, max_speed = 1,
         n_adj = 20) # ok
plot_adj('BW009', max_angle = 170, max_speed = 0.5) # reasonable

# BW010
out <- check_animal('BW010')
plot_adj('BW010', max_speed = 3.5) # ok
plot_adj('BW010', max_speed = 1, max_angle = 170) # ok
plot_adj('BW010', max_speed = 2, max_angle = 135) # ok

# BW012
out <- check_animal('BW012')
plot_adj('BW012', max_speed = 3)
plot_adj('BW012', max_speed = 1, max_angle = 170)
# location is ok
tel <- as.telemetry(d$tel[[which(d$animal == 'BW012')]])
i <- which(out$speed > 2 & out$angle > 170)
plot(tel, error = FALSE, type = 'l', xlim = c(-2600, 0),
     ylim = c(300, 2200), col = 'black')
plot(tel[i + -10:10, ],
     error = FALSE, type = 'l', col = 'blue', lwd = 2, add = TRUE)
plot(tel[i, ],
     error = FALSE, col = 'blue', pch = 19, add = TRUE)
plot(tel, add = TRUE)
rm(tel, i)

# BW013
out <- check_animal('BW013')
plot_adj('BW013', max_speed = 1, max_angle = 170) # uses linear features

# BW014 has a clear GPS error
out <- check_animal('BW014')
plot_adj('BW014', max_speed = 4)
flag_outlier(id = 'BW014', max_speed = 4, value = 1)
out <- check_animal('BW014')
plot_adj('BW014', max_speed = 2) # ok
plot_adj('BW014', max_speed = 0.5, min_speed = 0.6, max_angle = 170,
         n_adj = 30) # seems near a linear feature
# plot_adj('BW014', max_speed = 0.5, min_speed = 0.6, max_angle = 170,
#          map = TRUE) # moving along a river

# B027
out <- check_animal('BW027')
plot_adj('BW027', max_speed = 3, max_angle = 90) # not problematic

# BW028
out <- check_animal('BW028')
plot_adj('BW028', max_speed = 3) # not problematic

# B029
out <- check_animal('BW029')
plot_adj('BW029', max_speed = 2, max_angle = 175) # need a finer scale
plot_adj('BW029', max_speed = 2, max_angle = 175, n_adj = 2) # ok

# B031
out <- check_animal('BW031')
plot_adj('BW031', max_speed = 1, max_angle = 170) # need a finer scale
plot_adj('BW031', max_speed = 1, max_angle = 170, n_adj = 5) # ok

# B033: sharp and fast turns are near linear features
out <- check_animal('BW033')
plot_adj('BW033', max_speed = 1.5, max_angle = 90, n_adj = 5)

# BW044nex
out <- check_animal('BW044nex')
plot_adj('BW044nex', max_speed = 1, max_angle = 170)

# BW046nex
out <- check_animal('BW046nex')
plot_adj('BW046nex', max_speed = 1, max_angle = 175, min_speed = 3) # ok
plot_adj('BW046nex', max_speed = 1, min_speed = 3, # problematic
         max_angle = 135, min_angle = 175, reset_layout = FALSE)
i <- which(out$speed > 1 & out$speed < 3 & # likely hunting
             out$angle > 135 & out$angle < 175) + 0:5
points(location.lat ~ location.long,
       d$tel[[which(d$animal == 'BW046nex')]][i, ], col = 'blue', cex = 2)
d$tel[[which(d$animal == 'BW046nex')]][i, 'outlier'] <- 1
plot_adj('BW046nex', max_speed = 1, max_angle = 170) # ok
```

# *Cervus canadensis*

```{r cervus-canadensis, fig.pos='!H', fig.dim=c(10, 12)}
#' *NOTE:* assuming movement at ~< 0.25 m/s is realistic, even if angle >
#' 170 degrees since it's too hard to distinguish between movement and gps
#' error at that scale, and the models account for GPS error anyway (which
#' I assume to be 10 m since I don't have calibration data. Mike said most
#' collars have an error ~10 m anyway).

# E002
out <- check_animal('E002')
plot_adj('E002', max_speed = 0.4)
plot_adj('E002', max_angle = 170, max_speed = 0.2)

# E003
out <- check_animal('E003')
plot_adj('E003', max_speed = 0.5)

# E006
out <- check_animal('E006')
plot_adj('E006', max_speed = 0.8) # ok
plot_adj('E006', max_speed = 0.4, max_angle = 170) # outlier
#plot_adj('E006', max_speed = 0.4, max_angle = 170, map = TRUE) # see river
# dt is much larger: collar struggled to receive signal with tree cover
'hours' %#% out[which(out$speed > 0.4 & out$angle > 170) + (-4:4), 'dt']
'hours' %#% c(43282, 7254)
flag_outlier('E006', max_speed = 0.4, max_angle = 170, value = 1)
out <- check_animal('E006')
plot_adj('E006', max_speed = 0.2, max_angle = 170) # ok

# E007
out <- check_animal('E007')
plot_adj('E007', max_speed = 0.2, max_angle = 170) # ok

# E008
out <- check_animal('E008')
plot_adj('E008', max_speed = 0.3) # ok

# E011: last point is outlier (keeping previous points bc many 0 speeds)
out <- check_animal('E011')
plot_adj('E011', max_speed = 0.4, n_adj = 10)
# plot_adj('E011', max_speed = 0.4, n_adj = 10, map = TRUE)
plot_adj('E011', max_speed = 0.4, n_adj = 9, reset_layout = FALSE)
which(out$speed > 0.4) == nrow(out)
i_tel <- which(d$animal == 'E011')
i <- (nrow(d$tel[[i_tel]]) - 7):nrow(d$tel[[i_tel]]) # drop last 8 points
points(location.lat ~ location.long, d$tel[[i_tel]][i, ], col = 'blue',
       cex = 2, pch = 19)
d$tel[[i_tel]][i, 'outlier'] <- 1
points(location.lat ~ location.long, d$tel[[i_tel]][i, ],
       col = as.numeric(d$tel[[i_tel]]$outlier[i]) + 1, cex = 2)
layout(1)

out <- check_animal('E011')
tel <- as.telemetry(d$tel[[i_tel]], mark.rm = TRUE)
plot(tel[nrow(tel) + -10:0, ], error = FALSE, type = 'b',
     pch = as.character(0:9))
plot_adj('E011', max_speed = 0.25)

# E013
out <- check_animal('E013')
plot_adj('E013', max_speed = 0.2, max_angle = 170) # escaped something?

# E015 has multiple GPS malfunctions
out <- check_animal('E015')
plot_adj('E015', max_speed = 1, reset_layout = FALSE)
# can't filter easily because the outliers don't have the max velocity
i <- d %>%
  pull(tel) %>%
  nth(which(d$animal == 'E015')) %>%
  mutate(i = 1:n()) %>%
  slice(which(timestamp ==
                as.POSIXct(out[which(out$speed > 0.6)[2], 't'])) +
          0:1) %>%
  pull(i) %>%
  first() + 0:10
points(location.lat ~ location.long, cex = 2, lwd = 3, col = 'blue',
       filter(d, animal == 'E015')$tel[[1]][i, ])
lines(location.lat ~ location.long,
      filter(d, animal == 'E015')$tel[[1]][i, ], lwd = 6, col = 'red3')
d$tel[[which(d$animal == 'E015')]][i, 'outlier'] <- 1
out <- check_animal('E015')
plot_adj('E015', max_speed = 0.60) # track looks ok now

# E016
out <- check_animal('E016')
plot_adj('E016', max_speed = 0.2, max_angle = 170) # one outlier
plot_adj('E016', max_speed = 0.2, max_angle = 179) # use higher upper bound
flag_outlier('E016', max_speed = 0.2, max_angle = 179, value = 1)
out <- check_animal('E016')
plot_adj('E016', max_speed = 0.2, max_angle = 170) # ok

# E017
out <- check_animal('E017')
plot_adj('E017', max_speed = 0.6) # ok
plot_adj('E017', max_angle = 170, max_speed = 0.2)

# E019
out <- check_animal('E019') # many high speeds at angle > 170
plot_adj('E019', max_speed = 0.65, max_angle = 135) # some outliers
plot_adj('E019', max_speed = 0.65, max_angle = 160, reset_layout = FALSE)
i <- which(out$speed > 0.65 & out$angle > 160)[c(1, 3)] # some outliers
points(location.lat ~ location.long,
       d$tel[[which(d$animal == 'E019')]][i, ], cex = 2, col = 'blue')
d$tel[[which(d$animal == 'E019')]][i, 'outlier'] <- 1

out <- check_animal('E019') # check other high speeds with angle > 160
plot_adj('E019', max_speed = 0.4, max_angle = 135, reset_layout = FALSE)
tel <- d$tel[[which(d$animal == 'E019')]]
i <- which(out$speed > 0.4 & out$angle > 135)
points(location.lat ~ location.long, cex = 1.5, col = 'black',
       tel[! tel$outlier, ][i, ],
       pch = as.character(0:(length(i) - 1)))
'hour' %#% out[i, 'dt'] # consistent sampling

# 0 is likely an error (erring on the side of caution)
# 1 is consistent with other movement
# 2 is likely an error
# 3 is consistent with other movement
# 4 is likely an error
# 5 is realistic
# 6 is likely an error
# 7 is realistic
# 8 is likely an error
# 9 is likely an error
i <- i[1 + c(0, 2, 4, 6, 8, 9)]
i_id <- which(d$animal == 'E019')
d$tel[[i_id]][! d$tel[[i_id]]$outlier, ][i, 'outlier'] <- 1
# ensure correct locations were removed
out <- check_animal('E019') # check other high speeds with angle > 160
plot_adj('E019', max_speed = 0.4, max_angle = 135, reset_layout = FALSE)

# E023
out <- check_animal('E023')
plot_adj('E023', max_speed = 1) # ok
plot_adj('E023', max_speed = 0.6, min_speed = 1) # ok
plot_adj('E023', max_speed = 0.3, max_angle = 170) # ok

# E024
out <- check_animal('E024')
plot_adj('E024', max_speed = 0.4) # ok
plot_adj('E024', max_speed = 0.2, max_angle = 170) # ok

# E025
out <- check_animal('E025')
plot_adj('E025', max_speed = 0.3, max_angle = 170) # ok

# E026
#' this elk does not have an unusual amount of data
hist(map_int(filter(d, species == 'Cervus canadensis')$tel, nrow))
abline(v = nrow(d$tel[[which(d$animal == 'E026')]]), lwd = 2)
# d$tel[[which(d$animal == 'E026')]] %>%
#   filter(! outlier) %>%
#   sf::st_as_sf(coords = c('location.long', 'location.lat'),
#                crs = 4326) %>%
#   mapview(map.types = 'Esri.WorldImagery', col.regions = 'darkorange',
#           col = 'black', lwd = 2)
out <- check_animal('E026') # many sharp turns at high speeds
plot_adj('E026', max_speed = 0.6, max_angle = 135)
#plot_adj('E026', max_speed = 0.6, max_angle = 135, map = TRUE)# look north
flag_outlier('E026', max_speed = 0.6, max_angle = 135, value = 1)

out <- check_animal('E026')
plot_adj('E026', max_speed = 0.4, max_angle = 170) # NW one is ok
# plot_adj('E026', max_speed = 0.4, max_angle = 170, map = TRUE)
plot_adj('E026', max_speed = 0.42, max_angle = 170) # W one may also be ok
plot_adj('E026', max_speed = 0.42, max_angle = 175) # W one may also be ok
flag_outlier('E026', max_speed = 0.42, max_angle = 175, value = 1)

out <- check_animal('E026')
plot_adj('E026', max_speed = 0.35, max_angle = 170) # NW, middle, & NE ok?
plot_adj('E026', max_speed = 0.35, min_speed = 0.4, max_angle = 170,
         reset_layout = FALSE)
i <- with(out, which(speed > 0.35 & speed < 0.4 & angle > 160))[c(3:4)]
points(location.lat ~ location.long,
       filter(d$tel[[which(d$animal == 'E026')]], ! outlier)[i, ],
       cex = 2, col = 'blue', lwd = 2)
#' values in `i` assume outliers have been removed
i_tel <- which(d$animal == 'E026')
d$tel[[i_tel]][d$tel[[i_tel]]$outlier == 0, ][i, 'outlier'] <- 1

out <- check_animal('E026')
plot_adj('E026', max_speed = 0.30, max_angle = 160) # likely ok

#' using an error-informed, continuous-time movement model substantially
#' reduces the issues caused by GPS error. Additionally, there are so many
#' elk data (for each individual and the number of elk, too) that a few
#' values from a single individual will have little to no effect.

# E027
out <- check_animal('E027')
plot_adj('E027', max_speed = 0.3, max_angle = 170, n_adj = 4) # not all
plot_adj('E027', max_speed = 0.3, max_angle = 175, n_adj = 4,
         reset_layout = FALSE)
i <- with(out, which(speed > 0.3 & angle > 175))[c(2, 3, 4, 5)]
points(location.lat ~ location.long,
       d$tel[[which(d$animal == 'E027')]][i, ],
       cex = 2, col = 'blue', lwd = 2)
'hours' %#% out$dt[i]
d$tel[[which(d$animal == 'E027')]][i, 'outlier'] <- 2 # not fully sure

# E030
out <- check_animal('E030')
plot_adj('E030', max_speed = 0.45) # ok
plot_adj('E030', max_speed = 0.3, max_angle = 135) # generally ok
plot_adj('E030', max_speed = 0.3, max_angle = 170) # possibly problematic
flag_outlier('E030', max_speed = 0.3, max_angle = 170, value = 1)

out <- check_animal('E030')
plot_adj('E030', max_speed = 0.2, max_angle = 170) # ok

# E032
out <- check_animal('E032')
plot_adj('E030', max_speed = 0.3, max_angle = 170) # ok

# E033
out <- check_animal(id = 'E033')
plot_adj('E033', max_speed = 0.6)

# E036
out <- check_animal(id = 'E036')
# the next point is along the same line 
plot_adj('E036', max_speed = 0.7, reset_layout = FALSE)
points(location.lat ~ location.long, pch = 19,
       filter(d, animal == 'E036')$tel[[1]][0:1 + which(out$speed > 0.7),])
plot_adj('E036', max_speed = 0.6, reset_layout = FALSE) # ok

# E037
out <- check_animal(id = 'E037')
plot_adj(id = 'E037', max_speed = 0.3, max_angle = 170) # ok

# E040
out <- check_animal(id = 'E040')
plot_adj('E040', max_speed = 0.35, max_angle = 160) # ok

# E042: range shift with highly stationary behavior
out <- check_animal(id = 'E042')
tel <- as.telemetry(d$tel[[which(d$animal == 'E042')]])
range(tel$timestamp)
plot(tel)
# plot_adj(id = 'E042', many = TRUE, map = TRUE, max_speed = 0)
ctmm.guess(tel)

# E049
out <- check_animal(id = 'E049')
plot_adj('E049', max_speed = 0.4, max_angle = 170) # ok
plot_adj('E049', max_speed = 0.3, min_speed = 0.4, max_angle = 170) # ok

# E067
out <- check_animal(id = 'E067')
plot_adj('E067', max_speed = 0.6, max_angle = 170) # ok

# E068
out <- check_animal(id = 'E068')
plot_adj('E068', max_speed = 0.65, reset_layout = FALSE)
out[out$speed > 0.65, ] # 1st 2 points are right after capture (rownames)
points(location.lat ~ location.long,
       filter(d, animal == 'E068')$tel[[1]][1:2, ], col = 'blue', cex = 2)
d[d$animal == 'E068', ]$tel[[1]]$outlier[1:2] <- 1
out <- check_animal(id = 'E068')
plot_adj('E068', max_speed = 0.4, max_angle = 170)
# plot_adj('E068', max_speed = 0.4, max_angle = 170, map = TRUE)
flag_outlier('E068', max_speed = 0.4, max_angle = 170, value = 1)
out <- check_animal(id = 'E068')
plot_adj('E068', max_speed = 0.25, max_angle = 170) # ok

# E070
out <- check_animal('E070')
plot_adj('E070', max_speed = 0.4, max_angle = 135) # ok

# E074
out <- check_animal(id = 'E074')
plot_adj('E074', max_speed = 0.4, max_angle = 170) # ok

# E075
out <- check_animal(id = 'E075')
plot_adj('E075', max_speed = 0.3, max_angle = 170) # ok

# E079
out <- check_animal(id = 'E079')
plot_adj('E079', max_speed = 0.6) # ok

# E080
out <- check_animal(id = 'E080')
plot_adj('E080', max_speed = 0.8) # ok
plot_adj('E080', max_speed = 0.5, max_angle = 170, reset_layout = FALSE)
i_tel <- which(d$animal == 'E080')
i <- which(out$speed > 0.5 & out$angle > 170)[1:2]
points(location.lat ~ location.long,
       filter(d$tel[[i_tel]], ! outlier)[i, ], col = 'blue',
       cex = 2)
d$tel[[i_tel]][! d$tel[[i_tel]]$outlier, ]$outlier[i] <- 1
out <- check_animal(id = 'E080')
plot_adj('E080', max_speed = 0.5, max_angle = 170) # ok

# E081
out <- check_animal(id = 'E081')
plot_adj('E081', max_speed = 0.2, max_angle = 170) # ok

# E082
out <- check_animal(id = 'E082')
plot_adj('E082', max_speed = 0.2, max_angle = 170) # ok

# E083
out <- check_animal(id = 'E083')
plot_adj('E083', max_speed = 0.3, max_angle = 135) # ok

# E084 is not worth keeping
out <- check_animal(id = 'E084')
plot(location.lat ~ location.long, filter(d, animal == 'E084')$tel[[1]],
     type = 'b', pch = 19, col = '#00000010')
nrow(filter(d, animal == 'E084')$tel[[1]])
range(filter(d, animal == 'E084')$tel[[1]]$timestamp)
d <- filter(d, animal != 'E084')

# E085
out <- check_animal(id = 'E085')
plot_adj('E085', max_speed = 1) # definitely outliers (> 20 km!)
flag_outlier(id = 'E085', max_speed = 1, value = 1)

out <- check_animal(id = 'E085')
plot_adj('E085', max_speed = 0.6) # maybe ok
with(out[which(out$speed > 0.6) + -10:10, ], # likely outlier
     plot('hours' %#% dt, dt * speed / 1e3,
          col = as.numeric(-10:10 == 0) + 1,
          pch = 19, ylab = 'SLD (dt * min speed, km)'))
flag_outlier('E085', max_speed = 0.6, value = 1)

out <- check_animal(id = 'E085')
plot_adj('E085', max_speed = 0.2, max_angle = 135) # ok

# E096
out <- check_animal(id = 'E096')
plot_adj('E096', max_speed = 0.4, max_angle = 170) # ok

# E100
out <- check_animal(id = 'E100')
plot_adj('E100', max_speed = 0.3, max_angle = 170) # ok

# E101
out <- check_animal(id = 'E101')
plot_adj('E101', max_speed = 0.4, max_angle = 170) # ok

# E106
d$tel[[which(d$animal == 'E106')]] %>%
  group_by(gps.fix.type.raw) %>%
  summarise(hdop = mean(dop, na.rm = TRUE))
d$tel[[which(d$animal == 'E106')]] %>%
  group_by(gps.fix.type.raw) %>%
  summarise(hdop = mean(! is.na(dop), na.rm = TRUE))
out <- check_animal(id = 'E106')
plot_adj('E106', max_speed = 0.4, max_angle = 170) # large DOP values; ok
d$tel[[which(d$animal == 'E106')]] %>%
  as.telemetry() %>%
  data.frame() %>%
  slice(which(out$speed > 0.4 & out$angle > 170)) %>%
  pull(HDOP)
plot_adj('E106', max_speed = 0.3, max_angle = 170) # ok
plot_adj('E106', max_speed = 0.2, max_angle = 170) # ok

# E115
# collar is very error-prone in SW forest
out <- check_animal(id = 'E115')
# plot_adj('E115', many = TRUE, map = TRUE)
plot_adj('E115', max_speed = 0.4, max_angle = 170) # improbable
plot_adj('E115', max_speed = 0.4, max_angle = 170, n_adj = 50)#but possible
plot_adj('E115', max_speed = 0.4, max_angle = 175) # unlikely
slice(d$tel[[which(d$animal == 'E115')]],
      which(out$speed > 0.4 & out$angle > 175))$dop # low DOP
flag_outlier(id = 'E115', max_speed = 0.4, max_angle = 175, value = 2)

out <- check_animal('E115')
plot_adj('E115', max_speed = 0.3, max_angle = 170) # unlikely
flag_outlier('E115', max_speed = 0.3, max_angle = 170, value = 2)

out <- check_animal('E115')
plot_adj('E115', max_speed = 0.1, max_angle = 170, many = TRUE) # ok

# E116
out <- check_animal(id = 'E116')
plot_adj('E116', max_speed = 0.4, max_angle = 170)
flag_outlier(id = 'E116', max_speed = 0.4, max_angle = 170, value = 1)

out <- check_animal('E116')
plot_adj('E116', max_speed = 0.8)
flag_outlier(id = 'E116', max_speed = 0.8, value = 1)

out <- check_animal('E116')
plot_adj('E116', max_speed = 0.3, max_angle = 170) # improbable
plot_adj('E116', max_speed = 0.3, max_angle = 175) # but not impossible

# E117 has similar error patterns to E115
out <- check_animal(id = 'E117')
plot_adj('E117', max_speed = 0.7) # three outliers
# plot_adj('E117', max_speed = 0.7, map = TRUE)
flag_outlier(id = 'E117', max_speed = 0.7, value = 1)

out <- check_animal('E117')
#plot_adj('E117', many = TRUE, map = TRUE) # more errors when more forested
plot_adj('E117', max_speed = 0.4, max_angle = 170)

# plot_adj('E117', max_speed = 0.4, max_angle = 170, map = TRUE)
flag_outlier(id = 'E117', max_speed = 0.4, max_angle = 170, value = 1)

out <- check_animal('E117')
plot_adj('E117', max_speed = 0.3, max_angle = 170) # maybe not all
plot_adj('E117', max_speed = 0.3, max_angle = 175) # likely all
#plot_adj('E117', max_speed = 0.3, max_angle = 175, map = TRUE) # in forest
flag_outlier(id = 'E117', max_speed = 0.3, max_angle = 175, value = 1)

out <- check_animal('E117')
# still an excessive amount of sharp turns with high speeds (compare E118)
# dropping some of the data is better than dropping the whole animal
flag_outlier('E117', max_speed = 0.1, max_angle = 130, value = 2)
out <- check_animal('E117')

# E118 (compare with E117 above)
check_animal('E118', return_out = FALSE)

# E122
out <- check_animal(id = 'E122')
plot_adj('E122', max_speed = 0.45) # ok

# E125
out <- check_animal(id = 'E125')
plot_adj('E125', max_speed = 0.4, max_angle = 150)
flag_outlier('E125', max_speed = 0.4, max_angle = 150, value = 1)
out <- check_animal(id = 'E125') # ok

# E129
out <- check_animal(id = 'E129')
plot_adj('E129', max_speed = 0.6) # ok

# E131
out <- check_animal(id = 'E131')
plot_adj('E131', max_speed = 0.3, max_angle = 170) # N one is ok
plot_adj('E131', max_speed = 0.45, max_angle = 170) # outlier
flag_outlier('E131', max_speed = 0.45, max_angle = 170, value = 1)

out <- check_animal(id = 'E131') # ok

# E132
#' some high minimum speeds at long sampling intervals for this individual
out <- check_animal(id = 'E132', cap_dt = FALSE)
layout(matrix(c(0, 0, 0, 1), ncol = 2)) #
abline(v = 20, col = 'red')
abline(h = 0.08, col = 'red')
i <- which(out$dt > 20 %#% 'hours' & out$speed > 0.08)
out[i, ]
# small range shift at the daily scale within home range
plot_adj('E132', max_speed = 0.08, max_dt = 20 %#% 'hours')

# look at points near values with excessively high speeds
i_adj <- sapply(i, function(.i) .i + -10:10) %>%
  as.data.frame() %>% # convert matrix to a data frame
  pivot_longer(! c()) # use long format for a column of all values
tel <- as.telemetry(d$tel[[which(d$animal == 'E132')]])
adj <- left_join(slice(data.frame(out), i_adj$value),
                 slice(data.frame(tel), unique(i_adj$value)),
                 by = c('t')) %>%
  bind_cols(i_adj, .) %>%
  mutate(date = as.POSIXct(t),
         high_speed = speed > 0.08)
layout(matrix(1:4, ncol = 2))
for(g in seq(unique(i_adj$name))) {
  g_name <- unique(i_adj$name)[g]
  plot(y ~ x, filter(adj, name == g_name), type = 'l')
  points(y ~ x, data.frame(tel[i, ][g, ]), col = 'blue', pch = 19)
  rm(g, g_name)
} # ok
layout(1)

# large dt with minor "range shifts" ==> ok
plot('days' %#% dt ~ as.POSIXct(t), out[i_adj$value, ])
points('days' %#% dt ~ as.POSIXct(t), out[i, ], col = 'blue', pch = 19)

rm(i, i_adj, tel, adj)

# E134
out <- check_animal(id = 'E134')
plot_adj('E134', max_speed = 0.5, max_angle = 170) # ok

# E138
out <- check_animal(id = 'E138')
plot_adj('E138', max_speed = 0.8) # ok
plot_adj('E138', max_speed = 0.3, max_angle = 170) # ok

# E139
out <- check_animal(id = 'E139')
plot_adj('E139', max_speed = 0.3, max_angle = 170) # ok

# E140
out <- check_animal(id = 'E140')
plot_adj('E140', max_speed = 0.6)
flag_outlier('E140', max_speed = 0.6, value = 1)

out <- check_animal(id = 'E140') # ok

# E142
out <- check_animal(id = 'E142')
flag_outlier('E142', max_speed = 10, value = 1) # definitely an outlier

out <- check_animal(id = 'E142')
plot_adj('E142', max_speed = 0.8) # ok

# E143
out <- check_animal(id = 'E143')
plot_adj('E143', max_speed = 0.2, max_angle = 170) # ok

# E144
out <- check_animal(id = 'E144')
plot_adj('E144', max_speed = 0.3, max_angle = 170) # all ok

# E147
out <- check_animal(id = 'E147')
plot_adj('E147', max_speed = 0.3, max_angle = 170) # ok

# E161
out <- check_animal(id = 'E161')
plot_adj('E161', max_speed = 0.3, max_angle = 170) # ok

# E164: first point is definitely an outlier
out <- check_animal(id = 'E164')
tel <- d$tel[[which(d$animal == 'E164')]]
plot(location.lat ~ location.long, tel, type = 'b')
points(location.lat ~ location.long, tel[1, ], col = 'red', pch = 19)
rm(tel)
d$tel[[which(d$animal == 'E164')]]$outlier[1] <- 1

out <- check_animal(id = 'E164')
plot_adj('E164', max_speed = 0.3, max_angle = 160) # ok
rm(tel)

# E168
out <- check_animal(id = 'E168')
plot_adj('E168', max_speed = 0.3, max_angle = 170)
flag_outlier('E168', max_speed = 0.3, max_angle = 170, value = 1)

out <- check_animal(id = 'E168')
plot_adj('E168', max_speed = 0.2, max_angle = 170) # ok

# E170
out <- check_animal(id = 'E170')
plot_adj('E170', max_speed = 1)
plot_adj('E170', max_speed = 0.3, max_angle = 170) # ok

# E171
out <- check_animal(id = 'E171')
plot_adj('E171', max_speed = 0.6) # ok
plot_adj('E171', max_speed = 0.3, max_angle = 170) # ok

# E172
out <- check_animal(id = 'E172')
plot_adj('E172', max_speed = 0.3, max_angle = 150) # ok

# E178
out <- check_animal(id = 'E178')
plot_adj('E178', max_speed = 0.3, max_angle = 150) # ok

# E179
out <- check_animal(id = 'E179')
plot_adj('E179', max_speed = 0.8) # ok

# E181
out <- check_animal(id = 'E181')
plot_adj('E181', max_speed = 0.5, max_angle = 170) # ok

# E181
out <- check_animal(id = 'E182')
plot_adj('E182', max_speed = 0.3, max_angle = 170) # ok
```

# *Oreamnos americanus*

```{r oreamnos-americanus, fig.pos='!H', fig.dim=c(10, 12)}
d %>%
  filter(species == 'Oreamnos americanus') %>%
  unnest(tel) %>%
  pull(hdop) %>%
  hist(main = 'Goat HDOP before as.telemetry()')

d %>%
  filter(species == 'Oreamnos americanus') %>%
  unnest(tel) %>%
  pull(pdop) %>%
  hist(main = 'Goat PDOP before as.telemetry()')

d %>%
  filter(species == 'Oreamnos americanus') %>%
  mutate(tel = map(tel, \(.t) data.frame(as.telemetry(.t)))) %>%
  unnest(tel) %>%
  pull(HDOP) %>%
  hist(main = 'Goat HDOP after as.telemetry()')

# pdop is not in the telemetries
d %>%
  filter(species == 'Oreamnos americanus') %>%
  mutate(tel = map(tel, \(.t) data.frame(as.telemetry(.t)))) %>%
  unnest(tel) %>%
  colnames()

# substitute missing HDOP values with PDOP values
d <- d %>%
  mutate(
    tel = if_else(species == 'Oreamnos americanus',
                  map(tel, \(.t) {
                    .t %>%
                      mutate(hdop = if_else(is.na(hdop), pdop, hdop)) %>%
                      return()
                  }),
                  tel))

# ensure values are ok
all((d %>%
       filter(species == 'Oreamnos americanus') %>%
       unnest(tel) %>%
       pull(hdop)) ==
      (d %>%
         filter(species == 'Oreamnos americanus') %>%
         mutate(tel = map(tel, \(.t) data.frame(as.telemetry(.t)))) %>%
         unnest(tel) %>%
         pull(HDOP)))

# all goats have some extreme outliers in Berlin
filter(d, species == 'Oreamnos americanus') %>%
  unnest(tel) %>%
  ggplot() +
  facet_wrap(~ animal) +
  geom_hex(aes(location.long, location.lat), bins = 10) +
  geom_vline(xintercept = 0, color = 'darkorange') +
  scale_fill_continuous(limits = c(1, NA), trans = 'log10') +
  theme(legend.position = c(0.8, 0.15))

filter(d, species == 'Oreamnos americanus') %>%
  unnest(tel) %>%
  filter(location.long < 0) %>%
  pull(location.long) %>%
  hist()
abline(v = -119.8, col = 'darkorange')

# set all locations with long > 0 to outliers
d <-
  mutate(
    d,
    tel = if_else(species == 'Oreamnos americanus',
                  true =
                    map(tel, \(.d) {
                      mutate(
                        .d,
                        outlier = if_else(location.long > -119.8, 1, 0),
                        original_outliers = outlier == 1)
                    }),
                  false = tel))

# 30548
out <- check_animal(id = '30548')
hist(out$angle) # a lot of sharp turns at low speeds
plot_adj('30548', max_speed = 0.4) # ok
plot_adj('30548', max_speed = 0.3, max_angle = 135) # ok
# plot_adj('30548', max_speed = 0.3, max_angle = 135, map = TRUE) # ridges
plot_adj('30548', max_speed = 0.2, max_angle = 135) # ok
plot_adj('30548', max_speed = 0.1, max_angle = 170, n_adj = 3) # ok

# 30551
out <- check_animal(id = '30551') # another 124, many with speed = 0
# start and end are definitely not part of the goat's movement
d$tel[[which(d$animal == '30551')]] %>%
  filter(! outlier) %>%
  ggplot(aes(location.long, location.lat, col = as.POSIXct(timestamp))) +
  geom_path() +
  geom_point() +
  scale_color_viridis_c()
flag_outlier('30551', max_distance = 2e4, value = 1)

out <- check_animal(id = '30551') # last point is still a clear outlier
plot_adj('30551', max_speed = 0.6)
which(out$speed > 0.6) == nrow(out)
flag_outlier('30551', max_speed = 0.6, value = 1)

out <- check_animal(id = '30551') # all other speeds are reasonable
plot_adj('30551', max_speed = 0.4, n_adj = 20) # ok

# 30561
out <- check_animal(id = '30561')
# plot_adj('30561', max_distance = 15e3, map = TRUE)
flag_outlier('30561', max_distance = 15e3, value = 1)

out <- check_animal(id = '30561') # some high speeds at sharp turns
plot_adj('30561', max_speed = 0.2, max_angle = 160)
#plot_adj('30561', max_speed = 0.2, max_angle = 160, map = TRUE) # outliers
flag_outlier('30561', max_speed = 0.2, max_angle = 160, value = 1)

out <- check_animal(id = '30561') # speeds are reasonable
# plot_adj('30561', max_speed = 0, map = TRUE, many = TRUE) # ok

# 30567
out <- check_animal(id = '30567')
plot_adj('30567', max_speed = 0.3, max_angle = 160)
flag_outlier('30567', max_speed = 0.3, max_angle = 160, value = 1)

out <- check_animal(id = '30567')
plot_adj('30567', max_speed = 0.3) # ok
plot_adj('30567', max_speed = 0.2, max_angle = 160)
flag_outlier('30567', max_speed = 0.2, max_angle = 160, value = 1)

out <- check_animal(id = '30567')
plot_adj('30567', max_speed = 0.15, max_angle = 160) # NE is ok
plot_adj('30567', max_speed = 0.15, max_angle = 170)
plot_adj('30567', max_speed = 0.15, max_angle = 170, n_adj = 3)
#plot_adj('30567', max_speed = 0.15, max_angle = 170, n_adj = 3, map = TRUE)
flag_outlier('30567', max_speed = 0.15, max_angle = 170, value = 1)

out <- check_animal(id = '30567') # some possible outliers, but unclear
# plot_adj('30567', max_speed = 0, many = TRUE, map = TRUE)
plot_adj('30567', max_speed = 0.1, max_angle = 170) # ok
out[out$speed > 0.1 & out$angle > 170, ]
plot_adj('30567', max_speed = 0.1, max_angle = 170, max_dt = 2e4) # ok

# 30575
out <- check_animal(id = '30575')
plot_adj('30575', max_speed = 0.8)
flag_outlier('30575', max_speed = 0.8, value = 1)

out <- check_animal(id = '30575')
plot_adj('30575', max_speed = 0.5)
flag_outlier('30575', max_speed = 0.5, value = 1)

out <- check_animal(id = '30575')
plot_adj('30575', max_speed = 0.3, max_angle = 170)
flag_outlier('30575', max_speed = 0.3, max_angle = 170, value = 1)

out <- check_animal(id = '30575')
plot_adj('30575', max_speed = 0.2, max_angle = 170)
flag_outlier('30575', max_speed = 0.2, max_angle = 170, value = 1)

out <- check_animal(id = '30575')
plot_adj('30575', max_speed = 0.175, max_angle = 170)
flag_outlier('30575', max_speed = 0.175, max_angle = 170, value = 1)

out <- check_animal(id = '30575')
plot_adj('30575', max_speed = 0.15, max_angle = 170)
# plot_adj('30575', max_speed = 0.15, max_angle = 170, map = TRUE)
flag_outlier('30575', max_speed = 0.15, max_angle = 170, value = 1)

out <- check_animal(id = '30575')
plot_adj('30575', max_speed = 0.12, max_angle = 170)
plot_adj('30575', max_speed = 0.12, max_angle = 170, max_dt = 2e4)
flag_outlier('30575', max_speed = 0.12, max_angle = 170, max_dt = 2e4,
             value = 1)

out <- check_animal(id = '30575') # some points in forest, but speed is low
# plot_adj('30575', many = TRUE, map = TRUE)
plot(as.telemetry(d$tel[[which(d$animal == '30575')]], mark.rm = TRUE))

# 30599
out <- check_animal(id = '30599')
plot_adj(id = '30599', max_speed = 0.3, max_angle = 170)
flag_outlier(id = '30599', max_speed = 0.3, max_angle = 170, value = 1)

out <- check_animal(id = '30599')
plot_adj(id = '30599', max_speed = 0.2, max_angle = 170)
flag_outlier(id = '30599', max_speed = 0.2, max_angle = 170, value = 1)

out <- check_animal(id = '30599')
plot_adj(id = '30599', max_speed = 0.3)
flag_outlier(id = '30599', max_speed = 0.3, value = 1)

out <- check_animal(id = '30599')
plot_adj(id = '30599', max_speed = 0.15, max_angle = 170)
flag_outlier(id = '30599', max_speed = 0.15, max_angle = 170, value = 1)

out <- check_animal(id = '30599')
plot_adj(id = '30599', max_speed = 0.1, max_angle = 175)
out[out$angle > 175 & out$speed > 0.1, ]
plot_adj(id = '30599', max_speed = 0.1, max_angle = 175, min_dt = 2e4)
i <- which(out$speed > 0.1 & out$angle > 175 & out$dt < 2e4)
tel <- d$tel[[which(d$animal == '30599')]]
d$tel[[which(d$animal == '30599')]][! tel$outlier, ][i, 'outlier'] <- 1

out <- check_animal(id = '30599') # ok

# 30613
out <- check_animal(id = '30613')
plot_adj('30613', max_speed = 0.4, max_angle = 170)
# plot_adj('30613', max_speed = 0.4, max_angle = 170, map = TRUE)
flag_outlier('30613', max_speed = 0.4, max_angle = 170, value = 1)

out <- check_animal(id = '30613')
plot_adj('30613', max_speed = 0.2, max_angle = 170)
flag_outlier('30613', max_speed = 0.2, max_angle = 170, value = 1)

out <- check_animal(id = '30613')
plot_adj('30613', max_speed = 0.2, max_angle = 135)
# plot_adj('30613', max_speed = 0.2, max_angle = 135, map = TRUE)
flag_outlier('30613', max_speed = 0.2, max_angle = 135, value = 1)

out <- check_animal(id = '30613')
plot_adj('30613', max_speed = 0.12, max_angle = 170)
flag_outlier('30613', max_speed = 0.12, max_angle = 170, value = 1)

out <- check_animal(id = '30613')
plot_adj('30613', max_speed = 0.1, max_angle = 170) # ok
plot_adj('30613', max_speed = 0.1, max_angle = 150) # ok
plot_adj('30613', max_speed = 0.05, max_angle = 170) # ok

# 30636
out <- check_animal(id = '30636')
flag_outlier('30636', max_distance = 15e3, value = 1)

out <- check_animal(id = '30636')
plot_adj('30636', max_speed = 0.4)
flag_outlier('30636', max_speed = 0.4, value = 1)

out <- check_animal(id = '30636')
plot_adj('30636', max_speed = 0.1, max_angle = 170) # ok

# 30642
out <- check_animal(id = '30642')
plot_adj('30642', max_speed = 1)
flag_outlier('30642', max_speed = 1, value = 1)

out <- check_animal(id = '30642')
plot_adj('30642', max_speed = 0.3, max_angle = 170)
out[out$speed > 0.3 & out$angle > 170, ]
plot_adj('30642', max_speed = 0.3, max_angle = 170, max_dt = 8100)
flag_outlier('30642', max_speed = 0.3, max_angle = 170, max_dt = 8100,
             value = 1)

out <- check_animal(id = '30642')
plot_adj('30642', max_speed = 0.2, max_angle = 170, min_speed = 0.3)
i_tel <- which(d$animal == '30642')
i <- which(out$speed > 0.2 & out$speed < 0.3 & out$angle > 170)
d$tel[[i_tel]][! d$tel[[i_tel]]$outlier, ][i, 'outlier'] <- 1
plot_adj('30642', max_speed = 0.2, max_angle = 170, min_speed = 0.3) # ok

out <- check_animal(id = '30642') # ok

# 30648
out <- check_animal(id = '30648')
flag_outlier('30648', max_speed = 1, value = 1)

out <- check_animal(id = '30648')
plot_adj('30648', max_speed = 0.3, max_angle = 170)
flag_outlier('30648', max_speed = 0.3, max_angle = 170, value = 1)

out <- check_animal(id = '30648')
plot_adj('30648', max_speed = 0.2, max_angle = 170) # ok
plot_adj('30648', max_speed = 0.4, max_angle = 130)
flag_outlier('30648', max_speed = 0.4, max_angle = 130, value = 1)

out <- check_animal(id = '30648')
plot_adj('30648', max_speed = 0.2, max_angle = 170)
# plot_adj('30648', max_speed = 0.2, max_angle = 170, map = TRUE)
plot_adj('30648', max_speed = 0.2, max_angle = 175)
flag_outlier('30648', max_speed = 0.2, max_angle = 175, value = 1)

out <- check_animal(id = '30648')
plot_adj('30648', max_speed = 0.2, max_angle = 170) # ok
plot_adj('30648', max_speed = 0.1, max_angle = 175) # ok
```

# *Puma concolor*

```{r puma-concolor, fig.pos='!H', fig.dim=c(10, 12)}
# 246
out <- check_animal(id = '246')
plot_adj('246', max_speed = 3)
flag_outlier('246', max_speed = 3, value = 1)

out <- check_animal(id = '246')
plot_adj('246', max_speed = 0.5, n_adj = 3) # ok
plot_adj('246', max_speed = 0.2, max_angle = 170) # ok
# plot_adj('246', max_speed = 0.2, max_angle = 170, n_adj = 4, map = TRUE)
# most southern point may be an outlier, but it looks ok
plot_adj('246', max_speed = 0.2, max_angle = 177, min_angle = 178) # ok
# plot_adj('246', max_speed = 0.2, max_angle = 177, min_angle = 178,
#          n_adj = 4, map = TRUE)
i <- which(out$speed > 0.2 & out$angle > 177 & out$angle < 178) + -5:5
tel <- d$tel[[which(d$animal == '246')]]
tel <- tel[! tel$outlier, ]
plot(location.lat ~ location.long, tel[i, ], type = 'b',
     pch = as.character(seq(length(i))))
'hours' %#% out[i, 'dt']
rm(i, tel, cols)

# 272 has a clear speed outlier, but locations are ok
out <- check_animal(id = '272')
plot_adj('272', max_speed = 1) # data seem ok
plot_adj('272', max_speed = 0.2, max_angle = 135) # ok

# C1
out <- check_animal(id = 'C1')
plot_adj('C1', max_speed = 0.2, max_angle = 170) # ok

# C2
out <- check_animal(id = 'C2')
plot_adj('C2', max_speed = 0.2, max_angle = 170, n_adj = 20) # gps error
# plot_adj('C2', max_speed = 0.2, max_angle = 170, n_adj = 20, map = TRUE)
flag_outlier('C2', max_speed = 0.2, max_angle = 170, value = 1)

# C3
out <- check_animal(id = 'C3')
plot_adj('C3', max_speed = 1)
# plot_adj('C3', max_speed = 1, map = TRUE)
flag_outlier('C3', max_speed = 1, value = 1)
out <- check_animal(id = 'C3') # remaining data are ok
plot_adj('C3', max_speed = 0.4, max_angle = 170, n_adj = 5) # ok
plot_adj('C3', max_speed = 0.3, max_angle = 170, n_adj = 5) # ok

# C4
out <- check_animal(id = 'C4')
plot_adj('C4', max_speed = 0.64)
# plot_adj('C4', max_speed = 0.64, map = TRUE)
flag_outlier('C4', max_speed = 0.64, value = 1)
out <- check_animal(id = 'C4') # remaining data are ok
plot_adj('C4', max_speed = 0.2, max_angle = 170) # ok

# C5
out <- check_animal(id = 'C5')
plot_adj('C5', max_speed = 0.6) # ok
plot_adj('C5', max_speed = 0.4, max_angle = 170) # ok
plot_adj('C5', max_speed = 0.3, max_angle = 170) # ok

# C6
out <- check_animal(id = 'C6')
plot_adj('C6', max_speed = 0.4, max_angle = 170)
flag_outlier('C6', max_speed = 0.4, max_angle = 170, value = 1)
out <- check_animal(id = 'C6') # ok

# C7
out <- check_animal(id = 'C7')
plot_adj('C7', max_speed = 0.6, max_angle = 135)
flag_outlier('C7', max_speed = 0.6, max_angle = 135, value = 1)
out <- check_animal(id = 'C7') # ok
plot_adj('C7', max_speed = 0.4, max_angle = 135) # ok

# C8
out <- check_animal(id = 'C8')
plot_adj('C8', max_speed = 0.55) # ok
# plot_adj('C8', max_speed = 0.55, map = TRUE)
'hours' %#% out[which(out$speed > 0.55) + -5:5, 'dt']
plot_adj('C8', max_speed = 0.2, max_angle = 170) # ok

# C9
out <- check_animal(id = 'C10')
plot_adj('C10', max_speed = 0.2, max_angle = 170) # ok

# C10
out <- check_animal(id = 'C9')
plot_adj('C9', max_speed = 0.2, max_angle = 170) # ok

# C11
out <- check_animal(id = 'C11')
plot_adj('C11', max_speed = 2)
flag_outlier('C11', max_speed = 2, value = 1)

out <- check_animal(id = 'C11')
plot_adj('C11', max_speed = 0.4, max_angle = 170) # NE ok
# plot_adj('C11', max_speed = 0.4, max_angle = 175, map = TRUE) # outlier
'hours' %#% out[which(out$speed > 0.4 & out$angle > 175) + -3:3, 'dt']
flag_outlier('C11', max_speed = 0.4, max_angle = 175, value = 1)

out <- check_animal(id = 'C11')
plot_adj('C11', max_speed = 0.3, max_angle = 170) # NE ok

# C12
out <- check_animal(id = 'C12')
plot_adj('C12', max_speed = 0.8)
i <- which(out$speed > 0.8) + c(-1:0)
layout(t(0:1))
points(location.lat ~ location.long, d$tel[[which(d$animal == 'C12')]][i,],
       cex = 2, col = 'blue')
out$dt[i] # distance seems excessive for 2-hour sampling
d$tel[[which(d$animal == 'C12')]][i, 'outlier'] <- 2
out <- check_animal(id = 'C12')
plot_adj(id = 'C12', max_speed = 0.2, max_angle = 170) # ok
rm(i)

# c13
out <- check_animal(id = 'C13')
plot_adj('C13', max_speed = 0.3, max_angle = 170) # ok

# C14
out <- check_animal('C14')
plot_adj('C14', max_speed = 0.4, max_angle = 160) # ok

# C16
out <- check_animal('C16')
plot_adj('C16', max_speed = 0.3, max_angle = 170) # all outliers
flag_outlier('C16', max_speed = 0.3, max_angle = 170, value = 1)
out <- check_animal('C16') # ok

# C17
out <- check_animal('C17')
plot_adj('C17', max_speed = 0.2, max_angle = 170)
plot_adj('C17', max_speed = 0.35, max_angle = 170)
# one of the two is the last point of the telemetry
out[which((out$angle > 170 | is.na(out$angle)) & out$speed > 0.35), ]
i <- which(out$angle > 170 & out$speed > 0.35)
out[i, ]
# don't want to flag last point
d$tel[[which(d$animal == 'C17')]]$outlier[i] <- 1
out <- check_animal('C17')
plot_adj('C17', max_speed = 0.35, max_angle = 179) # ok

# C18
out <- check_animal('C18')
plot_adj('C18', max_speed = 0.6) # the two speeds are too similar
flag_outlier('C18', max_speed = 0.6, value = 2)
out <- check_animal('C18')
plot_adj('C18', max_speed = 0.2, max_angle = 170) # ok

# C19
out <- check_animal('C19')
plot_adj('C19', max_angle = 160, max_speed = 0.4) # NW is ok
plot_adj('C19', max_speed = 0.4, max_angle = 170)
flag_outlier('C19', max_speed = 0.4, max_angle = 170, value = 1)

out <- check_animal('C19') # ok

# C23
out <- check_animal('C23')
plot_adj('C23', max_speed = 0.3, max_angle = 160) # S is ok
plot_adj('C23', max_speed = 0.3, max_angle = 170) # unsure
plot_adj('C23', max_speed = 0.3, max_angle = 170, n_adj = 3)
out[which(out$angle > 170 & out$speed > 0.3) + -3:3, 'dt'] # too regular
# plot_adj('C23', max_speed = 0.3, max_angle = 170, map = TRUE) # outlier
flag_outlier('C23', max_speed = 0.3, max_angle = 170, value = 1)
out <- check_animal('C23')

# C28
out <- check_animal('C28')
plot_adj('C28', max_speed = 0.3, max_angle = 170)
flag_outlier('C28', max_speed = 0.3, max_angle = 170, value = 1)
out <- check_animal('C28')
plot_adj('C28', max_speed = 0.25, max_angle = 170) # ok

# C29
out <- check_animal('C29')
plot_adj('C29', max_speed = 0.5, max_angle = 160)
flag_outlier('C29', max_speed = 0.5, max_angle = 160, value = 1)
out <- check_animal('C29')

# C30
out <- check_animal('C30')
plot_adj('C30', max_speed = 0.6)
flag_outlier('C30', max_speed = 0.6, value = 1)

out <- check_animal('C30')
plot_adj('C30', max_speed = 0.3, max_angle = 170)
flag_outlier('C30', max_speed = 0.3, max_angle = 170, value = 1)

out <- check_animal('C30')
plot_adj('C30', max_speed = 0.3, max_angle = 135) # ok

# C31
out <- check_animal('C31')
plot_adj('C31', max_speed = 0.37, max_angle = 160)
out[which(out$speed > 0.37 & out$angle > 160) + -3:3, 'dt'] # too regular
# plot_adj('C31', max_speed = 0.37, max_angle = 160, map = TRUE) # outlier
flag_outlier('C31', max_speed = 0.37, max_angle = 160, value = 1)
out <- check_animal('C31') # ok
```

# *Rangifer tarandus caribou* (boreal)

```{r boreal-caribou, fig.pos='!H', fig.dim=c(10, 12)}
# HSCEK063
out <- check_animal('HSCEK063')
plot_adj('HSCEK063', max_speed = 0.6) # ok
plot_adj('HSCEK063', max_speed = 0.2, max_angle = 135) # ok

# SCEK007b
out <- check_animal('SCEK007b')
plot_adj('SCEK007b', max_speed = 0.6, n_adj = 25) # ok

# SCEK009
out <- check_animal('SCEK009')
plot_adj('SCEK009', max_speed = 0.4, max_angle = 135)
plot_adj('SCEK009', max_speed = 0.4, max_angle = 135, n_adj = 5)
plot_adj('SCEK009', max_speed = 0.4, max_angle = 135, n_adj = 30) # ok

# SCEK020b migrates
out <- check_animal('SCEK020b')
plot(as.telemetry(d$tel[[which(d$animal == 'SCEK020b')]]))
# plot_adj('SCEK020b', max_speed = 0, many = TRUE, map = TRUE)

# SCEK026b
out <- check_animal('SCEK026b')
plot_adj('SCEK026b', max_speed = 0.5, n_adj = 20) # ok

# SCEK033
out <- check_animal('SCEK033')
plot_adj('SCEK033', max_speed = 1) # ok

# SCEK036
out <- check_animal('SCEK036')
plot_adj('SCEK036', max_speed = 0.6) # ok

# SCEK044
out <- check_animal('SCEK044')
plot_adj('SCEK044', max_speed = 1) # ok

# SCEK047
out <- check_animal('SCEK047')
plot_adj('SCEK047', max_speed = 1) # ok

# SCEK050
out <- check_animal('SCEK050')
plot_adj('SCEK050', max_speed = 0.2, max_angle = 135) # ok

# SCEK079
out <- check_animal('SCEK079')
plot_adj('SCEK079', max_speed = 0.7) # ok

# SCEK079b
out <- check_animal('SCEK079b', cap_dt = FALSE)
plot_adj('SCEK079b', max_speed = 0.4) # ok
plot_adj('SCEK079b', max_speed = 0.15, max_angle = 170) # 2/4 outliers
plot_adj('SCEK079b', max_speed = 0.23, max_angle = 170) # possible outliers
flag_outlier('SCEK079b', max_speed = 0.23, max_angle = 170, value = 2)
out <- check_animal('SCEK079b', cap_dt = FALSE)

# SCEK114
out <- check_animal('SCEK114', cap_dt = FALSE)
plot_adj('SCEK114', max_speed = 0.4, max_angle = 160) # ok

# SCEK119
out <- check_animal('SCEK119', cap_dt = FALSE)
plot_adj('SCEK119', max_speed = 0.6, max_angle = 160) # outlier
flag_outlier('SCEK119', max_speed = 0.6, max_angle = 160, value = 2)
out <- check_animal('SCEK119', cap_dt = FALSE)

# SCEK133
out <- check_animal('SCEK133', cap_dt = FALSE)
plot_adj('SCEK133', max_speed = 0.9, max_angle = 170) # ok

# SCEK136
out <- check_animal('SCEK136', cap_dt = FALSE)
plot_adj('SCEK136', max_speed = 0.6, max_angle = 170) # unsure
plot_adj('SCEK136', max_speed = 0.6, max_angle = 170, n_adj = 40) # outlier
# plot_adj('SCEK136', max_speed = 0.6, max_angle = 170, n_adj = 40, map = T)
flag_outlier('SCEK136', max_speed = 0.6, max_angle = 170, value = 1)
out <- check_animal('SCEK136', cap_dt = FALSE)

# SCEK161b
out <- check_animal('SCEK161b', cap_dt = FALSE)
plot_adj('SCEK161b', max_speed = 0.4, max_angle = 135) # ok

# SCEK163
out <- check_animal('SCEK163', cap_dt = FALSE)
plot_adj('SCEK163', max_speed = 0.8) # ok

# SCEK169
out <- check_animal('SCEK169', cap_dt = FALSE)
hist(out$speed)
plot_adj('SCEK169', max_speed = 10, n_adj = 1)

# SCEK170
out <- check_animal('SCEK170', cap_dt = FALSE)
hist(out$speed)
plot_adj('SCEK170', max_speed = 150, n_adj = 2)
plot_adj('SCEK170', max_speed = 2)
flag_outlier('SCEK170', max_speed = 2, value = 1)

out <- check_animal('SCEK170', cap_dt = FALSE)
plot_adj('SCEK170', max_speed = 2)
flag_outlier('SCEK170', max_speed = 2, value = 1)

out <- check_animal('SCEK170', cap_dt = FALSE)
plot_adj('SCEK170', max_speed = 2)
flag_outlier('SCEK170', max_speed = 2, value = 1)

out <- check_animal('SCEK170', cap_dt = FALSE)
plot_adj('SCEK170', max_speed = 1)
flag_outlier('SCEK170', max_speed = 1, value = 1)

out <- check_animal('SCEK170', cap_dt = FALSE) # ok

# SCEK176
out <- check_animal('SCEK176')
plot_adj('SCEK176', max_speed = 0.6) # ok

# SCEK177
out <- check_animal('SCEK177')
plot_adj('SCEK177', max_speed = 0.2, max_angle = 170)
plot_adj('SCEK177', max_speed = 0.2, max_angle = 170, n_adj = 5)
flag_outlier('SCEK177', max_speed = 0.2, max_angle = 170, value = 1)
out <- check_animal('SCEK177', cap_dt = FALSE)
which.max(out$dt) # ok
nrow(out)

# SCEK181
out <- check_animal('SCEK181', cap_dt = FALSE)
hist(out$speed)
plot_adj('SCEK181', max_speed = 2, many = TRUE)
flag_outlier('SCEK181', max_speed = 2, value = 1)

out <- check_animal('SCEK181', cap_dt = FALSE)
plot_adj('SCEK181', max_speed = 2)
flag_outlier('SCEK181', max_speed = 2, value = 1)

out <- check_animal('SCEK181', cap_dt = FALSE)
plot_adj('SCEK181', max_speed = 2)
flag_outlier('SCEK181', max_speed = 2, value = 1)

out <- check_animal('SCEK181', cap_dt = FALSE)
plot_adj('SCEK181', max_speed = 0.5) # ok

# SCEK185
out <- check_animal('SCEK185')
plot_adj('SCEK185', max_speed = 0.3) # ok

# SCEK194
out <- check_animal('SCEK194')
plot_adj('SCEK194', max_speed = 0.15, max_angle = 170, # one outlier
         reset_layout = FALSE)
i <- which(out$angle > 170 & out$speed > 0.15)[2]
points(location.lat ~ location.long,
       d$tel[[which(d$animal == 'SCEK194')]][i, ],
       col = 'blue', cex = 2)
d$tel[[which(d$animal == 'SCEK194')]]$outlier[i] <- 1
out <- check_animal('SCEK194')

# SCEK201
out <- check_animal('SCEK201')
range('hours' %#% out$dt, na.rm = TRUE)
plot_adj('SCEK201', max_speed = 0.3, max_angle = 170, n = 50)
flag_outlier('SCEK201', max_speed = 0.3, max_angle = 170, value = 1)

out <- check_animal('SCEK201')
plot_adj('SCEK201', max_speed = 0.15, max_angle = 170, n_adj = 30)
flag_outlier('SCEK201', max_speed = 0.15, max_angle = 170, value = 1)

out <- check_animal('SCEK201')
plot_adj('SCEK201', max_speed = 0.1, max_angle = 170, n_adj = 30) # ok

# SCEK204
out <- check_animal('SCEK204')
plot_adj('SCEK204', max_angle = 170, max_speed = 0.1) # ok

# SCEK206
out <- check_animal('SCEK206')
plot_adj('SCEK206', max_distance = 15000)
out[out$distance > 15000, ] # last two points are outliers
nrow(out)
flag_outlier('SCEK206', max_distance = 15000, value = 1)
out <- check_animal('SCEK206')

# SCEK212
out <- check_animal('SCEK212')
plot_adj('SCEK212', max_speed = 0.25, max_angle = 150) # ok

# SCEK213
out <- check_animal('SCEK213')
plot_adj('SCEK213', max_speed = 0.25, max_angle = 150) # ok

# SCEK214
out <- check_animal('SCEK214')
plot_adj('SCEK214', max_speed = 0.25, max_angle = 150) # ok

# SCEK216
out <- check_animal('SCEK216')
plot_adj('SCEK216', max_speed = 0.4, max_angle = 135) # ok

# SCEK238
out <- check_animal('SCEK238')
plot_adj('SCEK238', max_speed = 0.5) # ok
```

# *Rangifer tarandus caribou* (southern mountain)

```{r sm-caribou, fig.pos='!H', fig.dim=c(10, 12)}
filter(d, dataset_name == 'Rangifer_tarandus_southern_mountain') %>%
  unnest(tel) %>%
  pull(hdop) %>%
  hist(main = 'Mountain caribou HDOP')

#' some rowns have both`hdop` and `dop` values
filter(d, dataset_name == 'Rangifer_tarandus_southern_mountain') %>%
  unnest(tel) %>%
  mutate(missing_hdop = is.na(hdop),
         missing_dop = is.na(dop)) %>%
  group_by(dataset_name, missing_hdop, missing_dop) %>%
  summarise(hdop = mean(hdop, na.rm = TRUE),
            pdop = mean(pdop, na.rm = TRUE),
            dop = mean(dop, na.rm = TRUE))

#' some have substantially different non-NA values for `hdop` and `dop`
#' assuming `ctmm` takes `hdop` and that `hdop` is better than `dop`
filter(d, dataset_name == 'Rangifer_tarandus_southern_mountain') %>%
  unnest(tel) %>%
  filter(! is.na(hdop) & ! is.na(dop)) %>%
  transmute(diff = hdop - dop) %>%
  pull(diff) %>%
  hist()

# check date ranges
filter(d, dataset_name == 'Rangifer_tarandus_southern_mountain') %>%
  unnest(tel) %>%
  group_by(animal) %>%
  summarise(range = diff(range(as.POSIXct(timestamp)))) %>%
  pull(range) %>%
  as.numeric() %>%
  hist(xlab = 'Days')

# 92_prepenned
ID <- '92_prepenned'
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.3)

# 101_prepenned
ID <- '101_prepenned'
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.5)
which(out$speed > 0.5)
d$tel[[which(d$animal == ID)]][1, 'outlier'] <- 1 # ran after collaring
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.2, max_angle = 170) # ok
plot_adj(ID, max_speed = 0.1, min_speed = 0.2, max_angle = 170) # ok

# 104_prepenned
ID <- '104_prepenned'
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.4) # ok

# 105_prepenned
ID <- '105_prepenned'
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.1, max_angle = 150) # ok

# 109_prepenned
ID <- '109_prepenned'
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.07, max_angle = 170) # ok

# 117_prepenned
ID <- '117_prepenned'
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.1) # ok

# 132_prepenned
ID <- '132_prepenned'
out <- check_animal(ID)
diff(range(as.POSIXct(out$t))) # only 15.5 days of data
plot_adj(ID, max_speed = 0.08) # ok
plot('hours' %#% out$dt)
plot_adj(ID, max_dt = 12 %#% 'hours') # ok

# 152_prepenned
ID <- '152_prepenned'
out <- check_animal(ID)
plot_adj(ID, max_speed = 0.1, max_angle = 135) # ok
```

# *Ursus arctos horribilis*

```{r ursus-arctos-horribilis, fig.pos='!H', fig.dim=c(10, 12)}
# 197
out <- check_animal('197')
plot_adj('197', max_speed = 0.3, max_angle = 170) # ok

# 242: first point is an outlier
out <- check_animal('242')
plot_adj('242', max_speed = 1, n_adj = 10)
which(out$speed > 1)
d$tel[[which(d$animal == '242')]][1, 'outlier'] <- 1
out <- check_animal('242')

# 259
out <- check_animal('259')
plot_adj('259', max_speed = 0.4, max_angle = 170, n_adj = 10)
flag_outlier('259', max_speed = 0.4, max_angle = 170, value = 1)
out <- check_animal('259')
plot_adj('259', max_speed = 0.2, max_angle = 135, n_adj = 5) # ok

# 260
out <- check_animal('260')
plot_adj('260', max_speed = 1, max_angle = 170, n_adj = 10)
flag_outlier('260', max_speed = 1, max_angle = 170, value = 1)
out <- check_animal('260')
plot_adj('260', max_speed = 0.4, max_angle = 170, n_adj = 10) # ok
plot_adj('260', max_speed = 0.3, max_angle = 170, n_adj = 10, # unsure
         min_speed = 0.4)
filter(data.frame(out), round(speed, 1) == 0.4 & angle > 170)
0.3509840 * 3645 / 1e3 # reasonable distance in an hour

# 262
out <- check_animal('262')
plot_adj('262', max_speed = 0.3, max_angle = 170) # ok
plot_adj('262', max_speed = 0.3, max_angle = 135) # ok

# 291
out <- check_animal('291')
plot_adj('291', max_speed = 0.3, max_angle = 170) # ok

# 292
out <- check_animal('292')
plot_adj('292', max_speed = 1.2, n_adj = 10)
flag_outlier(id = '292', max_speed = 1.2, value = 1)
out <- check_animal('292')
plot_adj('292', max_speed = 0.8) # ok
# plot_adj('292', max_speed = 0.8, map = TRUE) # moving along a river

# 297
out <- check_animal('297')
plot_adj('297', max_speed = 0.5) # ok

# 317
out <- check_animal('317')
plot_adj('317', max_speed = 1.5, reset_layout = FALSE) # previous = outlier
i <- which(out$speed > 1.5) - 1
points(location.lat ~ location.long,d$tel[[which(d$animal == '317')]][i, ],
       cex = 2, col = 'blue')
d$tel[[which(d$animal == '317')]][i, 'outlier'] <- 1
out <- check_animal('317')
plot_adj('317', max_speed = 0.8) # ok
plot_adj('317', max_speed = 0.3, max_angle = 170) # possible outliers...
plot_adj('317', max_speed = 0.3, max_angle = 160) # but common at 0.3 m/s

# 318
out <- check_animal('318')
plot_adj('318', max_speed = 0.5) # ok
plot_adj('318', max_speed = 0.4, min_speed = 0.5) # ok
```