### 2025-06-10 writing/appendix-a-cleaning-telemetry-data.Rmd
changed appendix to use elk

---
title: "\\Large The heat is on: Rising temperatures alter when, how, and where large terrestrial mammals move"
author:
  - name: Stefano Mezzini
    email: stefano.mezzini@ubc.ca
    institute: [braes, biol]
  - name: Chris H. Fleming
    institute: [ucf, smith]
  - name: Siobhan Darlington
    institute: [braes, biol]
  - name: Adam T. Ford
    institute: [braes, biol]
  - name: TJ Gooliaff
    institute: [forest]
  - name: Karen E. Hodges
    institute: [braes, biol]
  - name: Kirk Safford
    institute: bcp
  - name: Robert Serrouya
    institute: [braes, biol, biodiv]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - ucf: Department of Biology, University of Central Florida, Orlando, Florida 32816, United States.
  - smith: Smithsonian Conservation Biology Institute, National Zoological Park, 1500 Remount Rd., Front Royal, VA 22630, United States.
  - forest: British Columbia Ministry of Forests, Lands, Natural Resource Operations and Rural Development, Penticton, BC, Canada.
  - bcp: British Columbia Ministry of Environment and Parks, Penticton, BC, Canada.
  - biodiv: Wildlife Science Centre, Biodiversity Pathways, University of British Columbia Okanagan, Revelstoke, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'bc-mammals-temperature.bib'
csl: 'global-change-biology.csl'
fontsize: 12pt
# indent: true
header-includes:
    - \renewcommand{\figurename}{Figure A\!\!} # for "Figure Ax.
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font=scriptsize, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage{float} # to use fig.pos="H"
subparagraph: true # needed for \usepackage[small]{titlesec}
urlcolor: cyan
output:
  bookdown::pdf_document2:
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: true
---

```{r setup, include=FALSE}
# set chunk defaults
knitr::opts_chunk$set(echo = TRUE, out.width = '100%',
                      fig.align = 'center', cache = TRUE, warning = FALSE,
                      message = FALSE,
                      root.dir = 'H:/GitHub/bc-mammals-temperature')
```

\clearpage

In this appendix, we illustrate the cleaning procedure for the telemetry data. For the sake of brevity, we only show the code and outputs for some elk (*Cervus canadensis*); the full script is available at  [https://github.com/QuantitativeEcologyLab/bc-mammals-temperature/blob/main/analysis/02-remove-outliers.R](https://github.com/QuantitativeEcologyLab/bc-mammals-temperature/blob/main/analysis/02-remove-outliers.R). We suggest looking at the script since this appendix does not include the interactive maps that are generated by `plot_adj(..., map = TRUE)`. The interactive maps often show linear features (e.g., roads) and other geographic features (e.g., mountain ridges, rivers) that were useful to see during the data cleaning. Therefore, we have left the function calls in the document for completeness, but we have commented them out to prevent them from being executed. Below is an example of a map with some locations for elk E003, as generated by running\newline`plot_adj('E006', max_speed = 0.4, max_angle = 170, map = TRUE)`.

\singlespacing

```{r map-example, echo=FALSE, out.height='40%'}
knitr::include_graphics(path = 'H:/GitHub/bc-mammals-temperature/figures/E006-map-example.png',
                        rel_path = FALSE)
```

\doublespacing

We start by attaching the necessary packages and sourcing some custom functions. The custom functions are all available on GitHub in the `functions` folder of the repository located at [https://github.com/QuantitativeEcologyLab/bc-mammals-temperature](https://github.com/QuantitativeEcologyLab/bc-mammals-temperature). While this appendix only shows the removal of a single outlier point, the appendix details the approach we used to clean the telemetry data. We removed outlier points using `flag_outlier(..., value = 1)`, which leverages `ctmm::as.telemetry()` to remove rows that have `TRUE` or a positive, non-zero number in the `outlier` column. See the comments in each code chunk for additional detail and considerations about individual data points.

\singlespacing

<!-- \clearpage -->

\small

```{r packages}
library('dplyr')     # for data wrangling (mutate(), %>%, etc.)
library('tidyr')     # for data wrangling (nest(), unnest(), pivot_*, etc.)
library('purrr')     # for functional programming (map_***(), etc.)
library('ctmm')      # for movement models
library('lubridate') # for working with dates
library('mapview')   # for interactive maps
library('ggplot2')   # for fancy plots
theme_set(theme_bw())

# source custom functions
source('functions/outlier_plots.R') # to plot outlier diagnostic plots
source('functions/check_animal.R') # to run diagnostic plots
source('functions/plot_adj.R') # to plot n adjacent locations
source('functions/flag_outlier.R') # to mark outliers
source('functions/remove_outlier_flags.R') # to start over with an animal
```

\normalsize

Before cleaning the telemetry data, we do some basic checks:

\small

```{r checks}
# dataset checks ----
# some NA timestamps (not enough to make a difference)
readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(is.na(timestamp)) %>%
  group_by(dataset_name) %>%
  summarize(n = n())

# ensure all only have one column of HDOP
readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(! is.na(timestamp)) %>%
  mutate(outlier = if_else(outlier, 1, 0)) %>%
  filter(! is.na(hdop) + ! is.na(HDOP) > 1) %>%
  nrow()
```

<!-- to avoid splitting code chunk -->
\clearpage

```{r checks-2}
# should only have one column of DOP
readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(! is.na(timestamp)) %>%
  mutate(outlier = if_else(outlier, 1, 0)) %>%
  filter(! is.na(DOP) + ! is.na(gps.dop) > 1) %>%
  nrow()
```

\normalsize

We can now import the data...

\small

```{r import-data}
# import the full dataset ----
# import the dataset after dropping NAs
d <- readRDS('data/tracking-data/all-tracking-data-not-cleaned.rds') %>%
  filter(! is.na(timestamp)) %>%
  mutate(hdop = if_else(is.na(hdop), HDOP, hdop), # merge error columns
         dop = if_else(is.na(DOP), gps.dop, DOP),
         original_outliers = outlier,
         outlier = if_else(outlier, 1, 0)) %>% # make numeric
  select(! c(HDOP, DOP, gps.dop)) %>% # remove duplicate columns
  relocate(dop, .after = pdop) %>%
  nest(tel = ! c(species, dataset_name, animal))
d # nested tibble of tracking datasets
```

\normalsize

... and do some last few check before cleaning.

\small

```{r final-checks}
# ensure animal IDs are unique
sum(duplicated(d$animal))
```

\clearpage

```{r final-checks2}
# check which species have DOP values
d %>%
  unnest(tel) %>%
  group_by(species) %>%
  summarise(across(c(dop, hdop, pdop), ~ sum(.x, na.rm = TRUE))) %>%
  mutate(has_dop = dop + hdop + pdop > 0) %>%
  arrange(desc(has_dop))

#' see`finite`: some SM caribou have different values for hdop and dop
#' other datasets are ok
d %>%
  unnest(tel) %>%
  mutate(has_hdop = ! is.na(hdop),
         has_dop = ! is.na(dop),
         has_pdop = ! is.na(pdop)) %>%
  group_by(dataset_name, has_hdop, has_dop, has_pdop) %>%
  summarise(hdop = mean(hdop, na.rm = TRUE),
            pdop = mean(pdop, na.rm = TRUE),
            dop = mean(dop, na.rm = TRUE)) %>%
  mutate(finite = map_int(1:n(),
                          ~ sum(! is.nan(c(hdop[.x], pdop[.x], dop[.x])))))

d %>%
  filter(dataset_name == 'Rangifer_tarandus_southern_mountain') %>%
  unnest(tel) %>%
  filter(! is.na(hdop) & ! is.na(dop)) %>%
  plot(hdop ~ dop, ., main = paste('n =', nrow(.)))
abline(a = 0, b = 1, col = 'red', lwd = 2)
```

\normalsize

\doublespacing

Now that we have organized the full telemetry dataset, we can start cleaning the data. Locations are candidates for removal if they are far from the median location (i.e., large median deviation) or high minimum speed between locations. The minimum speed is calculated as the distance between locations divided by time between locations, since traveling in a straight line is the fastest possible path. To search for potential outliers, the `check_animal()` function creates five plots: The top plot shows the telemetry data with an appropriate two-point equidistant projection, and it can help detect any migrations or range shifts. The second plot, created by `ctmm::outlie(tel)` [where `tel` is a `telemetry` object, see @calabrese_ctmm_2016] shows the telemetry data using a two-point equidistant projection. Points are larger and redder if they are further from the median location, while segments are thicker and bluer if the minimum speed is large. Points that are near the median location and segments with small minimum speed may not be visible, while severe location errors will often be identifiable due to their very red points and thick blue segments (Fig. \@ref(fig:outlie)). The third plot, created by `plot(ctmm::outlie(tel))`, shows the minimum speed against median deviation. This plot is simply an alternative visualization of the second plot (without the spatial information). The fourth plot shows minimum speed against turning angle, since severe location errors will have high minimum speed and very sharp (i.e., large) turning angles (Fig. \@ref(fig:outlie)). The fifth plot shows minimum speed against sampling interval, which can be helpful to detect points that imply excessively high minimum speeds for long periods of time. Generally, animals tend to move faster (for long periods of time) when moving in a straight line (i.e., ballistic movement as opposed to exploratory movement).

```{r outlie, fig.width=6, fig.height=6, echo=FALSE, fig.cap="Example output from $\texttt{ctmm::outlie()}$ using randomly generated bivariate Gaussian locations (dots) and a purpousefully erroneous location. Dots are larger and redder if they are further from the median location, and segments are thicker and bluer if the estimated straight-line displacement is high. Since the erroneous location is so far from the median location, both median deviation and minimum speed are quite high, so it is immediately visible. Note that the turning angle is large because the data implies a near-180-degree turn."}
set.seed(2)
tel <- data.frame(timestamp = as.POSIXct(1:100),
                  long = (rnorm(100) / 1e3),
                  lat = (rnorm(100) / 1e3))
tel[30, 2:3] <- c(12, 1) / 1e3
tel <- as.telemetry(tel)
o <- outlie(tel)
rm(tel, o)
```

Ideally, the second plot (center-left) should show a good mix of points and lines of different sizes, with no points or lines clearly standing out. Plots 3-5 should show a decrease in $y$ values (vertical axis) as the $x$ values (horizontal axis) increase. The intensity of the decrease will depend on a series of factors, including the animal's species and age. See elk E006 below for an example of a telemetry dataset with a clear GPS error that exhibits

In the two bottom figures produced by `check_animal()`, where minimum speed is plotted against turning angle and sampling interval, high speeds should occur at smaller turning angles and finer sampling intervals. High speeds at large turning angles and sampling intervals are indicative of GPS errors, as in the case for animal E006. E002 and E003 do not show any clear issues. E002 has two points with high speeds and large turning angles that are worth checking, but plotting the adjacent telemetry points with `plot_tel()` indicates there are no issues. It is worth noting that all speed-based inferences are only appropriate if sampling frequency is fine enough to reconstruct at least an approximate movement path [@noonan_scale-insensitive_2019; @nathan_big-data_2022]. The example data in Fig. \@ref(fig:outlie) would be too coarse to make any inferences on minimum speed because there are no clear paths in the data: the animal seems to move around randomly (which is unsurprising, since the locations are random i.i.d. bivariate Gaussian draws).

<!-- \clearpage -->

\singlespacing

\small

```{r update-opts, include=FALSE}
# update chunk defaults
knitr::opts_chunk$set(echo = TRUE, out.width = '100%',
                      fig.align = 'center', cache = TRUE, warning = FALSE,
                      message = FALSE,
                      root.dir = 'H:/GitHub/bc-mammals-temperature')
```

```{r E002, fig.pos = '!H', fig.dim = c(10, 10)}
# E002
out <- check_animal('E002')
plot_adj('E002', max_speed = 0.4)
plot_adj('E002', max_angle = 170, max_speed = 0.2)
```

```{r E003, fig.pos = '!H', fig.dim = c(10, 10)}
# E003
out <- check_animal('E003')
plot_adj('E003', max_speed = 0.5)
```

```{r E006, fig.pos = '!H', fig.dim = c(10, 10)}
# E006
out <- check_animal('E006')
plot_adj('E006', max_speed = 0.8) # ok
plot_adj('E006', max_speed = 0.4, max_angle = 170) # outlier
plot_adj('E006', max_speed = 0.4, max_angle = 170, map = FALSE) # see river
# dt is much larger: collar struggled to receive signal with tree cover
'hours' %#% out[which(out$speed > 0.4 & out$angle > 170) + (-4:4), 'dt']
'hours' %#% c(43282, 7254) # convert seconds to hours
flag_outlier('E006', max_speed = 0.4, max_angle = 170, value = 1)
out <- check_animal('E006')
plot_adj('E006', max_speed = 0.2, max_angle = 170) # ok
```

```{r E011, fig.pos = '!H', fig.dim = c(10, 10)}
# E011: last point is outlier (keeping previous points bc many 0 speeds)
out <- check_animal('E011')
plot_adj('E011', max_speed = 0.4, n_adj = 10)
plot_adj('E011', max_speed = 0.4, n_adj = 10, map = FALSE)
plot_adj('E011', max_speed = 0.4, n_adj = 9, reset_layout = FALSE)
which(out$speed > 0.4) == nrow(out)
i_tel <- which(d$animal == 'E011')
i <- (nrow(d$tel[[i_tel]]) - 7):nrow(d$tel[[i_tel]]) # drop last 8 points
points(location.lat ~ location.long, d$tel[[i_tel]][i, ], col = 'blue',
       cex = 2, pch = 19)
d$tel[[i_tel]][i, 'outlier'] <- 1
points(location.lat ~ location.long, d$tel[[i_tel]][i, ],
       col = as.numeric(d$tel[[i_tel]]$outlier[i]) + 1, cex = 2)
layout(1)

out <- check_animal('E011')
tel <- as.telemetry(d$tel[[i_tel]], mark.rm = TRUE)
plot(tel[nrow(tel) + -10:0, ], error = FALSE, type = 'b',
     pch = as.character(0:9))
plot_adj('E011', max_speed = 0.25)
```

```{r E013, fig.pos = '!H', fig.dim = c(10, 10)}
# E013
out <- check_animal('E013')
plot_adj('E013', max_speed = 0.2, max_angle = 170) # escaped something?
```

```{r E015, fig.pos = '!H', fig.dim = c(10, 10)}
# E015 has multiple GPS malfunctions
out <- check_animal('E015')
plot_adj('E015', max_speed = 1, reset_layout = FALSE)
# can't filter easily because the outliers don't have the max velocity
i <- d %>%
  pull(tel) %>%
  nth(which(d$animal == 'E015')) %>%
  mutate(i = 1:n()) %>%
  slice(which(timestamp ==
                as.POSIXct(out[which(out$speed > 0.6)[2], 't'])) +
          0:1) %>%
  pull(i) %>%
  first() + 0:10
points(location.lat ~ location.long, cex = 2, lwd = 3, col = 'blue',
       filter(d, animal == 'E015')$tel[[1]][i, ])
lines(location.lat ~ location.long,
      filter(d, animal == 'E015')$tel[[1]][i, ], lwd = 6, col = 'red3')
d$tel[[which(d$animal == 'E015')]][i, 'outlier'] <- 1
out <- check_animal('E015')
plot_adj('E015', max_speed = 0.60) # track looks ok now
```

```{r E016, fig.pos = '!H', fig.dim = c(10, 10)}
# E016
out <- check_animal('E016')
plot_adj('E016', max_speed = 0.2, max_angle = 170) # one outlier
plot_adj('E016', max_speed = 0.2, max_angle = 179) # use higher upper bound
flag_outlier('E016', max_speed = 0.2, max_angle = 179, value = 1)
out <- check_animal('E016')
plot_adj('E016', max_speed = 0.2, max_angle = 170) # ok
```

```{r E017, fig.pos = '!H', fig.dim = c(10, 10)}
# E017
out <- check_animal('E017')
plot_adj('E017', max_speed = 0.6) # ok
plot_adj('E017', max_angle = 170, max_speed = 0.2)
```

```{r E019, fig.pos = '!H', fig.dim = c(10, 10)}
# E019
out <- check_animal('E019') # many high speeds at angle > 170
plot_adj('E019', max_speed = 0.65, max_angle = 135) # some outliers
plot_adj('E019', max_speed = 0.65, max_angle = 160, reset_layout = FALSE)
i <- which(out$speed > 0.65 & out$angle > 160)[c(1, 3)] # some outliers
points(location.lat ~ location.long,
       d$tel[[which(d$animal == 'E019')]][i, ], cex = 2, col = 'blue')
d$tel[[which(d$animal == 'E019')]][i, 'outlier'] <- 1

out <- check_animal('E019') # check other high speeds with angle > 160
plot_adj('E019', max_speed = 0.4, max_angle = 135, reset_layout = FALSE)
tel <- d$tel[[which(d$animal == 'E019')]]
i <- which(out$speed > 0.4 & out$angle > 135)
points(location.lat ~ location.long, cex = 1.5, col = 'black',
       tel[! tel$outlier, ][i, ],
       pch = as.character(0:(length(i) - 1)))
'hour' %#% out[i, 'dt'] # consistent sampling

# 0 is likely an error (erring on the side of caution)
# 1 is consistent with other movement
# 2 is likely an error
# 3 is consistent with other movement
# 4 is likely an error
# 5 is realistic
# 6 is likely an error
# 7 is realistic
# 8 is likely an error
# 9 is likely an error
i <- i[1 + c(0, 2, 4, 6, 8, 9)]
i_id <- which(d$animal == 'E019')
d$tel[[i_id]][! d$tel[[i_id]]$outlier, ][i, 'outlier'] <- 1
# ensure correct locations were removed
out <- check_animal('E019') # check other high speeds with angle > 160
plot_adj('E019', max_speed = 0.4, max_angle = 135, reset_layout = FALSE)
```

\clearpage

# References